---
title: extRatum package example
author: Nikos Patias
date: 'Oct 8, 2020'
slug: extratum-package-example
categories: []
tags: []
comments: no
showcomments: yes
showpagemeta: yes
---

# This notebook provides an example of `extRatum` package

In this example, I demonstrate the use of `extRatum` package.

I use openly available geospatial data to calculate the area covered by water bodies in each National Park in England.

```{r}
#devtools::install("C:/Users/User/Desktop/extRatum")
library(extRatum, quietly = TRUE)
library(sf, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(tmap, quietly = TRUE)
```

## Read data

First, I read in the boundaries of National Parks in England, published by Natural England.

The data downloaded from: https://naturalengland-defra.opendata.arcgis.com/datasets/d333c7529754444894e2d7f5044d1bbf_0.


```{r}
# read in the national parks boundaries
national_parks <- st_read("lake_data/National_Parks_(England)/National_Parks__England____Natural_England.shp")

# plot the boundaries using an interactive map
tmap_mode("view")

tm_shape(national_parks) +
  tm_borders()
```


Then I read in the water bodies boundaries in England published by the Environment Agency.

The data downloaded from: https://data.gov.uk/dataset/33dcd836-3813-4233-a3ca-856358312415/wfd-lake-water-bodies-cycle-1


```{r}
# read in the water bodies boundaries
water_bodies <- st_read("lake_data/EA_WFDLakeWaterBodiesCycle1_SHP_Full/data/WFD_Lake_Water_Bodies_Cycle_1.shp")

# plot the boundaries using an interactive map
tmap_mode("view")

tm_shape(water_bodies) +
  tm_borders()
```

## Run the function

Finally, it is time to run the  `areal_calc()` function to calculate the area covered by water bodies in each National Park in England. 

Note that we have to pass a planar coordinate system, so the algorithm be able to calculate areas. In this example I use British National Grid.



```{r}
# run the function from extRatum package
water_coverage <-
  areal_calc(water_bodies,
             national_parks,
             unique_id_code = "FID",
             crs = "epsg:27700")


```

The output of the function gives us the total area covered by water bodies in each National Park,the total area of each National Park and the ratio between the two.

Note that the areas are expressed in sqkm.

Given that everything is measured in sqkm, the ratio represents what is the % of area covered by water bodies by sqkm. In this way, we have a relative measure that can be compared across all National Parks and is independent of their size.


```{r}
head(water_coverage)

```

## Create maps

Finally, I append the data calculated to the original table of National Parks.

```{r}
# perform a join of tables 
national_parks_v2 <- left_join(national_parks, water_coverage, by = "FID")

```


We can now create two maps.

One that shows the total area covered by water bodies

```{r}
# create a choropleth map
tm_shape(national_parks_v2) +
  tm_fill("AreaCovered", style = "fisher", palette = "Blues", alpha = 0.6, id="NAME", title="Area covered by water bodies (in km2)") 
```

And a second, that shows the ratio of water bodies to the total are of each National Park.

```{r}
# create a choropleth map
tm_shape(national_parks_v2) +
  tm_fill("Ratio", style = "fisher", palette = "Blues", alpha = 0.6, id="NAME", title="Ration of water bodies to total area") 
```